##### Edges

An edge is the following:

  - Connects a drug to a target gene

```{r}
edges <- full_table %>%
  dplyr::filter(sig == TRUE
                #,
                # cell == "HUVEC"
  ) %>%
  dplyr::select(pert_iname, t_gn_sym, cell) %>%
  tidyr::separate_rows(t_gn_sym, sep = ";") %>%
  dplyr::mutate(t_gn_sym = stringr::str_trim(t_gn_sym)) %>%
  magrittr::set_colnames(c("source", "target", "cell"))

moa_drugs <- full_table %>%
  dplyr::filter(sig == TRUE) %>%
  dplyr::select(pert_iname, MOAss) %>%
  tidyr::separate_rows(MOAss, sep = ";") %>%
  dplyr::mutate(MOAss = stringr::str_trim(MOAss))  %>%
  set_colnames(c("pert_iname", "MOA"))%>%
  dplyr::filter(!is.na(MOA)) %>%
  dplyr::distinct(.keep_all=T)


drugsThatAreMtorInhibitors <- full_table %>%
  dplyr::filter(sig == TRUE) %>%
  dplyr::filter(grepl("MTOR", MOAss)) %>%
  dplyr::select(pert_iname) %>%
  dplyr::distinct() %>%
  dplyr::mutate(isMTORInhibitor = TRUE)


nodes <- data.frame(id = c(unique(edges$source),
                           unique(edges$target)),
                    type = c(rep("drug", length(unique(edges$source))),
                             rep("gene", length(unique(edges$target)))
                    )
) %>%
  dplyr::left_join(drugsThatAreMtorInhibitors, by = c("id"="pert_iname")) %>%
  dplyr::mutate(isMTORInhibitor = case_when(
    is.na(isMTORInhibitor) ~ FALSE,
    TRUE ~ TRUE
  ))%>%
  left_join(moa_drugs, by = c("id"="pert_iname"), multiple = "all")
```




### Significant Results

```{r}
sigResults <- queries_against_lincs2 %>%
  lapply(function(x){
    x@result %>%
      dplyr::filter(!is.na(t_gn_sym),
                    trend == "up",
                    WTCS_FDR < 0.05,
                    NCSct > 0.8) %>%
      dplyr::arrange(desc(NCSct))
  })

lapply(sigResults, function(x){x%>%nrow()})

ueries_against_lincs2$macrophage@result %>% dplyr::filter(!is.na(t_gn_sym), trend == "up", WTCS_FDR < 0.05) %>% View
```

### Try to look for cell type agnostic drugs

```{r}
queries_all <- sigResults %>%
  bind_rows(.id = "celltype")

queries_all2 <- queries_all %>%
  dplyr::group_by(celltype, pert_iname) %>%
  dplyr::rename(query_celltype = celltype) %>%
  dplyr::summarise(
    n=n(),
    mean_NCS_across_sig_LINCS2_cell_lines=mean(NCSct),
    mean_FDR_across_sig_LINCS2_cell_lines = mean(WTCS_FDR)
  ) %>%
  dplyr::rename(n_sig_LINCS2_cell_lines = n) %>%
  dplyr::arrange(pert_iname, desc(n_sig_LINCS2_cell_lines)) %>%
  dplyr::group_by(pert_iname) %>%
  dplyr::mutate(sum_ranking_purposes = sum(mean_NCS_across_sig_LINCS2_cell_lines)) %>%
  dplyr::arrange(desc(sum_ranking_purposes))  %>%
  left_join(lincs_pert_info2 %>% as.data.frame %>%
              dplyr::select(pert_iname, MOA, t_gn_sym, indication_class, action_type, max_phase ) %>% dplyr::distinct(.keep_all=TRUE))

View(queries_all2)

queries_all2 %>% saveRDS(here("output", "main_cns_result.rds"))

```

### Filter endothelial only

```{r}
endo_results <- queries_all2 %>%
  dplyr::filter(query_celltype == "endothelial")
```



#### Functional enrichment analysis

```{r}
drugs_hepatocyte <-query_against_customdb_hepatocyte2$PPF_vs_Veh_NC_0hr %>%
  result %>%
  as.data.frame %>%
  dplyr::filter(WTCS_FDR < 0.05) %>%
  dplyr::pull(pert) %>%
  unique
# Gene ontology
go_tea <- tsea_dup_hyperG(drugs = drugs_hepatocyte,
                          universe="Default",
                          type = "GO",
                          ont = "MF",
                          pvalueCutoff = 0.05,
                          pAdjustMethod = "BH",
                          qvalueCutoff = 0.1,
                          minGSSize = 10,
                          maxGSSize = 500)

result(go_tea) %>% View

dtnetplot(drugs = drugs(go_tea),
          set = "GO:0004674",
          ont = "MF")


```

- KEGG

```{r}
kegg_tea  <- tsea_dup_hyperG(drugs = drugs_hepatocyte,
                             universe="Default",
                             type = "KEGG",
                             pvalueCutoff = 0.5,
                             pAdjustMethod = "BH",
                             qvalueCutoff = 0.1,
                             minGSSize = 10,
                             maxGSSize = 500)

result(kegg_tea) %>% View()
```

- Reactome

```{r}
reactome_tea  <- tsea_dup_hyperG(drugs = drugs_hepatocyte,
                                 type = "Reactome",
                                 pvalueCutoff = 0.5,
                                 pAdjustMethod = "BH",
                                 qvalueCutoff = 0.5,
                                 readable=TRUE)

result(reactome_tea) %>% View
```


### DEPRECATED - Control Cohort

#### Version 1 

```{r}
control_drug_names <- nesdata$Control_full %>%
  split(x = ., f = .$drug) %>%
  names()

control_drugs <- nesdata$Control_full %>%
  split(x = ., f = .$drug) %>%
  lapply(function(x){
    x %>% 
      dplyr::select(Estimate, pv, Somamer) %>%
      dplyr::mutate(Score = (1 - pv) * Estimate) %>%
      dplyr::select(Somamer, Score) %>%
      tibble::column_to_rownames("Somamer") 
  }) %>%
  bind_cols() %>%
  set_colnames(paste0(control_drug_names, 
                      "__",
                      "Control",
                      "__",
                      "Control"))

build_custom_db(
  control_drugs, 
  h5file = file.path(output_dir, "control_drugs.h5"))

control_drugs_db_path <- file.path(output_dir, "control_drugs.h5")
file.exists(control_drugs_db_path)
```

#### Version 3 - ala Volcano Plot

```{r}
control_drugs_v2 <- nesdata$Control_full %>%
  split(x = ., f = .$drug) %>%
  lapply(function(x){
    x %>% 
      dplyr::select(Estimate, pv, Somamer) %>%
      dplyr::mutate(Score = Estimate) %>%
      dplyr::select(Somamer, Score) %>%
      tibble::column_to_rownames("Somamer") 
  }) %>%
  bind_cols() %>%
  set_colnames(paste0(control_drug_names, 
                      "__",
                      "Control",
                      "__",
                      "Control"))

build_custom_db(
  control_drugs_v2, 
  h5file = file.path(output_dir, "control_drugs_v2.h5"))

control_drugs_db_path_v2 <- file.path(output_dir, "control_drugs_v2.h5")
file.exists(control_drugs_db_path_v2)
```

#### Version 3 - Only use fold change

```{r}
control_drugs_v3 <- nesdata$Control_full %>%
  split(x = ., f = .$drug) %>%
  lapply(function(x){
    x %>% 
      dplyr::select(Estimate, pv, Somamer) %>%
      dplyr::mutate(Score = -log10(pv)*Estimate) %>%
      dplyr::select(Somamer, Score) %>%
      tibble::column_to_rownames("Somamer") 
  }) %>%
  bind_cols() %>%
  set_colnames(paste0(control_drug_names, 
                      "__",
                      "Control",
                      "__",
                      "Control"))

build_custom_db(
  control_drugs_v3, 
  h5file = file.path(output_dir, "control_drugs_v3.h5"))

control_drugs_db_path_v3 <- file.path(output_dir, "control_drugs_v3.h5")
file.exists(control_drugs_db_path_v3)
```

```{r, eval=FALSE}
# sigs <- list(
#   dep_anx = readr::read_csv(file.path(nesda_path, "NESDA_per_somamer_meds_Dep_Anx.csv")) ,
#   dep = readr::read_csv(file.path(nesda_path, "NESDA_per_somamer_meds_Depression.csv"))
# ) %>%
#   lapply(function(x){
#     split(x=., f=.$drug)
#   }) %>%
#   # Subset to the drugs which are also present in the 
#   # Control cohort
#   lapply(function(x){
#     x[names(x) %in% control_drug_names]
#   })


```


### DEPRECATED - Import in Dep_Anx signature

```{r}
dep_anx_query <- readr::read_csv(file.path(nesda_path, "NESDA_per_somamer_meds_Dep_Anx.csv")) %>%
  split(x=., f=.$drug)

# Subset to the drugs which are also present in the Control cohort. 
dep_anx_query <- dep_anx_query[names(dep_anx_query) %in% control_drug_names]

# Order Somamers by Score for preparation of 
# Up and Down sets
dep_anx_query %<>%
  lapply(function(x){
    x %>%
      dplyr::select(Somamer, Estimate, pv) %>%
      dplyr::mutate(Score = (1 - pv) * Estimate) %>%
      dplyr::select(Somamer, Score) %>%
      dplyr::arrange(Score) %>%
      tibble::column_to_rownames("Somamer") 
  })

dep_anx_query %>% View

dep_anx_query_up <- dep_anx_query %>%
  lapply(function(x){
    x %>%
      dplyr::filter(Score > 0) %>%
      rownames()
  })

dep_anx_query_down <- dep_anx_query %>%
  lapply(function(x){
    x %>%
      dplyr::filter(Score < 0) %>%
      rownames()
  })
```
#   for(drug in names(dep_anx_query_up)){
#    qsig_preps[[db]][[drug]] <- 
#      qSig(
#        query = list(upset = dep_anx_query_up[[drug]],
#                     downset = dep_anx_query_down[[drug]]),
#                     gess_method = "LINCS",
#                     refdb = db_paths[[db]])
#      
#    
#    queries[[db]][[drug]] <-
#      gess_lincs(
#        qSig = qsig_preps[[db]][[drug]],
#        sortby = "NCS",
#        tau = FALSE,
#        cmp_annot_tb = NULL,
#        cmp_name_col = NULL,
#        #Need or it fails
#        workers = 2
#      )
#    results_queries[[db]][[drug]] <- 
#      result(queries[[db]][[drug]] )
#    
#   }
# }


```{r}
qsig_test <- qSig(
  query = list(upset = dep_anx_query_up$A02BA02,
               downset = dep_anx_query_down$A02BA02),
                  gess_method = "LINCS", 
                  refdb = control_drugs_db_path)

test_query <- gess_lincs(qSig = qsig_test,
                  sortby = "NCS",
                  tau = FALSE,
                  cmp_annot_tb = NULL,
                  cmp_name_col = NULL, #Need or it fails
                  workers = 2)

result(test_query) %>% View()
```
## Old 

```{r}

rank_res$Anxiety %>%
  melt() %>%
  dplyr::rename(method = variable,
                rank = value) %>%
  ggplot(aes(x = rank, y = method)) +
  geom_violin(alpha=0.3, aes(fill = method)) +
  scale_fill_manual(values = pals::glasbey(n=3)) +
  geom_boxplot(width = 0.2, fill = "white") +
  facet_wrap( ~ group)

rank_res$Control %>%
  melt() %>% 
  set_colnames(c("Method", "Rank"))  %>%
  ggplot(aes(x = Rank, y = Method)) +
  geom_boxplot() +
  geom_violin(alpha=0.3)

rank_res$Anxiety %>%
  melt() %>%
  dplyr::rename(method = variable,
                rank = value) %>%
  ggplot(aes(x = rank, y = method)) +
  geom_violin(alpha=0.3, aes(fill = method)) +
  scale_fill_manual(values = pals::glasbey(n=3)) +
  geom_boxplot(width = 0.2, fill = "white") +
  facet_wrap( ~ group)

ggviolin(df, x = "dose", y = "len", fill = "dose",
         palette = c("#00AFBB", "#E7B800", "#FC4E07"),
         add = "boxplot", add.params = list(fill = "white"))+
  stat_compare_means(comparisons = my_comparisons, label = "p.signif")+ # Add significance levels
  stat_compare_means(label.y = 50)    


rank_res %>% 
  lapply(function(x){
    apply(x, 2, sum)
  })

rank_res %>% saveRDS(file.path(output_dir, "rank_res.rds"))

cs_ranks$volcano$Anxiety
    
#     
#     for(d in names(results_queries[[scoringmethod]])){
#       cs_ranks[[scoringmethod]][[d]] <- 
#         results_queries[[scoringmethod]][[d]] %>%
#         dplyr::arrange(desc(NCS)) %>%
#         dplyr::mutate(rank = 1:nrow(.)) %>%
#         dplyr::filter(pert == d) %>%
#         dplyr::pull(rank)
#     }
#     
# }

# for(scoringmethod in names(results_queries)){
#     cs_ranks[[scoringmethod]] <- list()
#     for(d in names(results_queries[[scoringmethod]])){
#       cs_ranks[[scoringmethod]][[d]] <- 
#         results_queries[[scoringmethod]][[d]] %>%
#         dplyr::arrange(desc(NCS)) %>%
#         dplyr::mutate(rank = 1:nrow(.)) %>%
#         dplyr::filter(pert == d) %>%
#         dplyr::pull(rank)
#     }
#     
# }

cs_ranks %<>%
  lapply(function(x){
    x %>%
      unlist() %>%
  as.data.frame() %>%
  set_colnames("rank") 
  })
  

rank_results <- cs_ranks %>% bind_cols() %>%
  set_colnames(names(results_queries)) 

rank_results %>% saveRDS(file.path(output_dir, "rank_results.rds"))

apply(rank_results, 2, sum)
```



