---
title: "Scratch: Create internal CMap-like database from NESDA data"
author: "Nhi Hin"
date: "2023-05-03"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Load Packages

```{r message=FALSE, warning=FALSE}
library(Seurat)
library(dplyr)
library(magrittr)
library(readr)
library(tibble)
library(ggplot2)
library(SingleCellExperiment)
library(scater)
library(here)
library(rhdf5)
library(signatureSearch)
library(signatureSearchData)
library(HDF5Array)
library(ExperimentHub)
library(cowplot)
theme_set(theme_cowplot())
```

## Output Directory

```{r}
output_dir <- here("output", "_scratch_nesda_Rmd")

ifelse(!dir.exists(output_dir),
       dir.create(output_dir), 
       FALSE)
```


## Objective

1. Load NESDA data and get it into a form where we can use it for our internal database of CMap-like datasets. 

2. Answer the question of which scoring mechanism is best. 



## Load Helper functions

```{r}
source(here("code", "helper_functions.R"))
```


## Import Data

- Updated to only read in the full datasets for each cohort. 

```{r}
nesda_path <- file.path("/mnt/atlas/work/general/external/studies/NESDA")

nesda_names <- list.files(nesda_path,
                          pattern = "*_full.csv",
                          full.names = FALSE) %>%
  gsub(x = .,
       pattern = ".csv|NESDA_per_somamer_meds_",
       replacement = "")

nesdata <- list.files(nesda_path,
                      pattern = "*_full.csv",
                      full.names = TRUE) %>%
  lapply(readr::read_csv) %>%
  set_names(nesda_names)

nesdata %>% lapply(dim)

# nesdata$healthy <- NULL
# healthy appears same as $Control
```

## Import Somamer to Protein Mappings

```{r}
somaMap <- readRDS("/mnt/atlas/general_archive/data/target_selection/cycle_5/data_in/somalogic_annotations/Somalogic-Human_7K_Annotations-Uniprot_Centric.rds")
View(somaMap)
```

## Initial Exploration

### Number of Somamers Per Drugs Per Cohort

```{r}
table_ndrugsPerSomamerPerCohort <- nesdata %>% 
  lapply(function(x){
    x %>% dplyr::group_by(drug) %>%
      dplyr::summarise(n=n())
  }) %>%
  bind_rows(.id = "cohort") 
table_ndrugsPerSomamerPerCohort %>%
  write_csv(file.path(output_dir,
                      "table_ndrugsPerSomamerPerCohort.csv"))
  
table_ndrugsPerCohort <- table_ndrugsPerSomamerPerCohort %>%
  dplyr::group_by(cohort) %>%
  dplyr::summarize(n=n()) 
table_ndrugsPerCohort %>%
  write_csv(file.path(output_dir, "table_ndrugsPerCohort.csv"))
```

### Drug Overlap Between Cohorts

- The majority of drugs are only in the Anxiety-Depression Cohort, or shared between this Cohort and the Healthy and Control cohorts. 
- Also a sizable number of drugs ~15? that are shared between all Cohorts except Bipolar. 
- Bipolar only has one drug. 

```{r}
table_drugsPerCohort <- table_ndrugsPerSomamerPerCohort %>% 
  dplyr::group_by(drug) %>%
  dplyr::summarize(cohort = list(cohort))
View(table_drugsPerCohort)

plot_overlapDrugsPerCohort <- table_drugsPerCohort %>%
  ggplot(aes(x = cohort))+
  geom_bar() +
  ggupset::scale_x_upset()
plot_overlapDrugsPerCohort
plot_overlapDrugsPerCohort %>%
  ggsave(
    file.path(output_dir, "plot_overlapDrugsPerCohort.pdf"),
    plot = .,
    units = "in",
    width = 14,
    height = 8
  )
```

- Get overlaps as a list

```{r}
table_drugsPerCohort2 <- table_ndrugsPerSomamerPerCohort %>% 
  dplyr::group_by(cohort) %>%
  dplyr::summarize(drug = list(drug)) 

drugsPerCohort_l <- table_drugsPerCohort2$drug %>%
  as.list() %>%
  set_names(table_drugsPerCohort2$cohort)

drugsPerCohort_overlap <- overlapGroups(drugsPerCohort_l)
```


### Fold Change of Somamers Between Drugs & Between Cohorts

```{r}
tables_fc_per_somamer <- nesdata %>%
  bind_rows(.id = "cohort") %>%
  dplyr::arrange(Somamer) %>%
  split(x=., f = .$Somamer) %>%
  lapply(function(x){
    x %>%
      dplyr::group_by(drug) %>%
      dplyr::arrange(drug)
  })

tables_fc <- nesdata %>%
  lapply(function(x) {
    x %>%
      dplyr::group_by(drug) %>%
      dplyr::summarise(
        mean_fc = mean(Estimate),
        sd_fc = sd(Estimate),
        min_fc = min(Estimate),
        max_fc = max(Estimate),
        med_fc = median(Estimate)
      )
  })
  
```



## Prepare Matrix for SignatureSearch DB

- Initially, we will test this with the Control subset of patients. 

- Create various versions of the DB:

1. Connectivity Score is `(1 - pv) * Estimate`.
2. Connectivity Score is `Estimate`. This seems a bit fishy though considering that the fold changes are all quite subtle - leaving out p-value information seems not great. At the same time, fold changes should be directly comparable across drugs 
3. Connectivity Score is -log10(p) * Estimate. This is more similar to that of a Volcano Plot.  The potential disadvantage here is that small p-values could skew the overall result more than the v1 method. 

### Control Cohort

#### Version 1 

```{r}
control_drug_names <- nesdata$Control_full %>%
  split(x = ., f = .$drug) %>%
  names()

control_drugs <- nesdata$Control_full %>%
  split(x = ., f = .$drug) %>%
  lapply(function(x){
    x %>% 
      dplyr::select(Estimate, pv, Somamer) %>%
      dplyr::mutate(Score = (1 - pv) * Estimate) %>%
      dplyr::select(Somamer, Score) %>%
      tibble::column_to_rownames("Somamer") 
  }) %>%
  bind_cols() %>%
  set_colnames(paste0(control_drug_names, 
                      "__",
                      "Control",
                      "__",
                      "Control"))

build_custom_db(
  control_drugs, 
  h5file = file.path(output_dir, "control_drugs.h5"))

control_drugs_db_path <- file.path(output_dir, "control_drugs.h5")
file.exists(control_drugs_db_path)
```

#### Version 3 - ala Volcano Plot

```{r}
control_drugs_v2 <- nesdata$Control_full %>%
  split(x = ., f = .$drug) %>%
  lapply(function(x){
    x %>% 
      dplyr::select(Estimate, pv, Somamer) %>%
      dplyr::mutate(Score = Estimate) %>%
      dplyr::select(Somamer, Score) %>%
      tibble::column_to_rownames("Somamer") 
  }) %>%
  bind_cols() %>%
  set_colnames(paste0(control_drug_names, 
                      "__",
                      "Control",
                      "__",
                      "Control"))

build_custom_db(
  control_drugs_v2, 
  h5file = file.path(output_dir, "control_drugs_v2.h5"))

control_drugs_db_path_v2 <- file.path(output_dir, "control_drugs_v2.h5")
file.exists(control_drugs_db_path_v2)
```

#### Version 3 - Only use fold change

```{r}
control_drugs_v3 <- nesdata$Control_full %>%
  split(x = ., f = .$drug) %>%
  lapply(function(x){
    x %>% 
      dplyr::select(Estimate, pv, Somamer) %>%
      dplyr::mutate(Score = -log10(pv)*Estimate) %>%
      dplyr::select(Somamer, Score) %>%
      tibble::column_to_rownames("Somamer") 
  }) %>%
  bind_cols() %>%
  set_colnames(paste0(control_drug_names, 
                      "__",
                      "Control",
                      "__",
                      "Control"))

build_custom_db(
  control_drugs_v3, 
  h5file = file.path(output_dir, "control_drugs_v3.h5"))

control_drugs_db_path_v3 <- file.path(output_dir, "control_drugs_v3.h5")
file.exists(control_drugs_db_path_v3)
```

### Database Paths

```{r}
db_paths <- list(
  rick = control_drugs_db_path,
  fc_only = control_drugs_db_path_v2,
  volcano = control_drugs_db_path_v3
)
```


## Test by querying Signatures of Other Cohorts against Control 

### Import Signatures & Split to Drug-Per-Dataframe

```{r}
sigs <- list(
  dep_anx = readr::read_csv(file.path(nesda_path, "NESDA_per_somamer_meds_Dep_Anx.csv")) ,
  dep = readr::read_csv(file.path(nesda_path, "NESDA_per_somamer_meds_Depression.csv"))
) %>%
  lapply(function(x){
    split(x=., f=.$drug)
  }) %>%
  # Subset to the drugs which are also present in the 
  # Control cohort
  lapply(function(x){
    x[names(x) %in% control_drug_names]
  })


```


### Import in Dep_Anx signature

```{r}
dep_anx_query <- readr::read_csv(file.path(nesda_path, "NESDA_per_somamer_meds_Dep_Anx.csv")) %>%
  split(x=., f=.$drug)

# Subset to the drugs which are also present in the Control cohort. 
dep_anx_query <- dep_anx_query[names(dep_anx_query) %in% control_drug_names]

# Order Somamers by Score for preparation of 
# Up and Down sets
dep_anx_query %<>%
  lapply(function(x){
    x %>%
      dplyr::select(Somamer, Estimate, pv) %>%
      dplyr::mutate(Score = (1 - pv) * Estimate) %>%
      dplyr::select(Somamer, Score) %>%
      dplyr::arrange(Score) %>%
      tibble::column_to_rownames("Somamer") 
  })

dep_anx_query %>% View

dep_anx_query_up <- dep_anx_query %>%
  lapply(function(x){
    x %>%
      dplyr::filter(Score > 0) %>%
      rownames()
  })

dep_anx_query_down <- dep_anx_query %>%
  lapply(function(x){
    x %>%
      dplyr::filter(Score < 0) %>%
      rownames()
  })
```

### Perform Query

- Below, 

```{r}
qsig_preps <- list()
queries <- list()
results_queries <- list()

for(db in names(db_paths)){
  qsig_preps[[db]] <- list()
  queries[[db]] <- list()
  results_queries[[db]] <- list()
  
  for(drug in names(dep_anx_query_up)){
   qsig_preps[[db]][[drug]] <- 
     qSig(
       query = list(upset = dep_anx_query_up[[drug]],
                    downset = dep_anx_query_down[[drug]]),
                    gess_method = "LINCS",
                    refdb = db_paths[[db]])
     
   
   queries[[db]][[drug]] <-
     gess_lincs(
       qSig = qsig_preps[[db]][[drug]],
       sortby = "NCS",
       tau = FALSE,
       cmp_annot_tb = NULL,
       cmp_name_col = NULL,
       #Need or it fails
       workers = 2
     )
   results_queries[[db]][[drug]] <- 
     result(queries[[db]][[drug]] )
   
  }
}

queries %>% saveRDS(file.path(output_dir, "queries.rds"))
results_queries %>% saveRDS(file.path(output_dir, "results_queries.rds"))
```

### Rankings

```{r}
cs_ranks <- list()

for(scoringmethod in names(results_queries)){
    cs_ranks[[scoringmethod]] <- list()
    for(d in names(results_queries[[scoringmethod]])){
      cs_ranks[[scoringmethod]][[d]] <- 
        results_queries[[scoringmethod]][[d]] %>%
        dplyr::arrange(desc(NCS)) %>%
        dplyr::mutate(rank = 1:nrow(.)) %>%
        dplyr::filter(pert == d) %>%
        dplyr::pull(rank)
    }
    
}

cs_ranks %<>%
  lapply(function(x){
    x %>%
      unlist() %>%
  as.data.frame() %>%
  set_colnames("rank") 
  })
  

rank_results <- cs_ranks %>% bind_cols() %>%
  set_colnames(names(results_queries)) 

rank_results %>% saveRDS(file.path(output_dir, "rank_results.rds"))

apply(rank_results, 2, sum)
```



```{r}
qsig_test <- qSig(
  query = list(upset = dep_anx_query_up$A02BA02,
               downset = dep_anx_query_down$A02BA02),
                  gess_method = "LINCS", 
                  refdb = control_drugs_db_path)

test_query <- gess_lincs(qSig = qsig_test,
                  sortby = "NCS",
                  tau = FALSE,
                  cmp_annot_tb = NULL,
                  cmp_name_col = NULL, #Need or it fails
                  workers = 2)

result(test_query) %>% View()
```
