---
title: "scratch_disgenet"
author: "Nhi Hin"
date: "2023-03-20"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Load Packages

```{r}
library(disgenet2r)
library(Seurat)
library(dplyr)
library(magrittr)
library(readr)
library(tibble)
library(ggplot2)
library(SingleCellExperiment)
library(scater)
library(here)
library(rhdf5)
library(signatureSearch)
library(signatureSearchData)
library(HDF5Array)
library(ExperimentHub)
```

## API Key

```{r}
Sys.setenv(DISGENET_API_KEY = "030b2e02ff356ed135085de8c366cbbacb3bf9a4")
```

## Summary of this notebook

- Given results table from running CMap through *signatureSearch*, we need a way to interpret the results in terms of enriched indications. 
- In this notebook, we perform over-representation analysis using DisGeNet's disease-gene associations. 

## Import data

### CMap Results (PPF vs Veh, Endothelial cells, vs. HUVEC LINCS2)

See `prepare_6021_data_for_querying_LINCS2.Rmd` for details. 

```{r}
full_table <- readRDS(here("data", "full_table_endo.rds"))
colnames(full_table)
```

## Neaten up table 

- Retain relevant columns
- Separate out gene symbols, one on each row. 

**Note**: Below, here we are focusing on the HUVEC cell line as this is most comparable to endothelial cells from the transcriptlomics data. 

```{r}
full_table_tidy <- full_table %>%
  dplyr::filter(cell == "HUVEC") %>%
  dplyr::select(pert, pert_iname, 
                starts_with("WTC"),
                starts_with("NCS"),
                MOAss,
                t_gn_sym,
                sig)   %>%
  tidyr::separate_rows(t_gn_sym, sep = ";") %>%
  dplyr::mutate(t_gn_sym =  stringr::str_trim(t_gn_sym)) 
nrow(full_table_tidy)

head(full_table_tidy)
```

- Number of gene targets per drug:

```{r}
full_table_tidy %>%
  dplyr::distinct(pert_iname, t_gn_sym) %>%
  dplyr::group_by(pert_iname) %>%
  dplyr::summarise(n=n()) %>%
  dplyr::pull(n) %>%
  hist(breaks = 50)
```

-Benoit: what kinds of drugs have 20> targets

```{r}
drugswith20targets <- full_table_tidy %>%
  dplyr::distinct(pert_iname, t_gn_sym) %>%
  dplyr::group_by(pert_iname) %>%
  dplyr::summarise(n=n()) %>%
  dplyr::filter(n >= 20)


```

## Filter sig results, extract gene symbols

```{r}
sig_res <- full_table_tidy %>%
  dplyr::filter(sig == TRUE) 
  

sig_res %>%
  dplyr::distinct(pert_iname, t_gn_sym) %>% 
  dplyr::group_by(pert_iname) %>%
  dplyr::summarise(n=n()) %>% .$n %>%
  hist(breaks=50)

  sig_res <- sig_res%>% .$t_gn_sym %>% unique()
sig_res
length(sig_res)
```



## Enrichment (over-representation) analysis

- This will simply be performed with `disease_enrichment` function already implemented in DisGeNet's R package. 

```{r}
res_enrich <- disease_enrichment(entities = sig_res)

summaryPlot <- plot( res_enrich, class = "Enrichment", count =3,  cutoff= 0.05, nchars=70)

summaryPlot %>% ggsave(here("output", "summaryPlot_disgenet.pdf"),
                       . ,width = 11, height = 20, units = "in")

res_table <- res_enrich@qresult %>%
  dplyr::mutate(sig = FDR <= 0.05)
View(res_table)
res_table$sig %>% table

sig_table <- res_table %>%
  dplyr::filter(sig == TRUE)

nrow(sig_table)
```


```{r}
# Summary plot of number of disease classes in sig results
summaryPlot2 <- sig_table %>%
  tidyr::separate_rows(disease_class_name, sep = ";") %>%
  dplyr::filter(!is.na(disease_class_name)) %>%
  dplyr::mutate(disease_class_name = stringr::str_trim(disease_class_name)) %>%
  dplyr::distinct(Description, disease_class_name) %>%
  dplyr::group_by(disease_class_name) %>%
  dplyr::summarise(n=n()) %>%
  dplyr::arrange(desc(n)) %>%
  ggplot(aes(x = n, y = forcats::fct_reorder(disease_class_name, n)),
         fill = disease_class_name) +
  geom_bar(stat = "identity") +
  labs(x = "Number of significant diseases in each disease class",
       y = "Disease Class from DisGeNet Database")+
  cowplot::theme_cowplot()
summaryPlot2
ggsave(here("output", "summaryPlot2_disgenet.pdf"), plot = summaryPlot2,
       height = 9, width = 11, units = "in")
View(sig_table)
```


```{r}
res_table %>% readr::write_csv(here("output", "DisGeNet_TargetGene_ORA_All.csv"))
sig_table %>% readr::write_csv(here("output", "DisGeNet_TargetGene_ORA.csv"))
```

## Interpretation of gene-drug associations

- Create table of gene - drug pairs from the genes which were from significant drugs (based on CMap analysis). 

```{r}
genes_df <- data.frame(gene = sig_res)

disease_df <- res_table %>%
  dplyr::select(ID, Description, shared_symbol) %>%
  dplyr::rename(gene = shared_symbol) %>%
  tidyr::separate_rows(gene, sep = ";") %>%
  dplyr::distinct(.keep_all=T)

# Inner join as we only want to keep those pairs 
# which contain genes from the significant genes 
# list in genes_df. 
gene_disease_pairs <- dplyr::inner_join(
  genes_df,
  disease_df,
  by = "gene"
) %>% dplyr::select(-Description) %>%
  dplyr::distinct(.keep_all=TRUE)
View(gene_disease_pairs)
nrow(gene_disease_pairs)
```

- Retrieve evidence

```{r}
# This will be used to store the results from 
# disGeNET's gene2evidence function. 
all_gene_evidence_df <- list()

# For gene and drug pairs which do not have 
# evidence, skip them and continue looping
# Takes >30 mins to run. Need to 
# figure out how to parallelize at some point. 
for (i in 1:nrow(gene_disease_pairs)){
    tryCatch({
        tmp <- gene2evidence(
    gene = gene_disease_pairs$gene[[i]],
    vocabulary = "HGNC",
    disease = gene_disease_pairs$ID[[i]],
    database = "ALL",
    score = c(0.3, 1)
  )
  all_gene_evidence_df[[i]] <- disgenet2r::extract(tmp)
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}
```

- Condense down results into single data.frame

```{r}
all_gene_evidence_df2 <- all_gene_evidence_df %>%
  bind_rows()

all_gene_evidence_df2 %>% saveRDS(here("data", "all_gene_evidence_df2.rds"))
```




## Network visualizations
- Edge and node tables will be exported below, for visualization in Gephi. 

### Subnetwork - Nervous system diseases

```{r}
sig_table$disease_class_name %>% unique
```

```{r}
nervousSystemDiseases <- sig_table %>% 
  tidyr::separate_rows(disease_class_name, sep = ";") %>%
  dplyr::filter(!is.na(disease_class_name)) %>%
  dplyr::mutate(disease_class_name = stringr::str_trim(disease_class_name)) %>%
  # Make a CNS-themed network with the following 
  # disease classes.
  dplyr::filter(
  disease_class_name == "Nervous System Diseases" #|
    #disease_class_name == "Mental Disorders"
) %>%
  dplyr::select(Description, shared_symbol) %>%
  tidyr::separate_rows(shared_symbol, sep = ";") 
```

Let's see if we can create a network from this. 

#### Edges

edges between
- disease and target gene (DisGeNet information)
- drug and target gene (CMap information)

```{r}
drugGeneEdges <- full_table_tidy %>% 
  dplyr::filter(sig == TRUE) %>%
  dplyr::select(pert_iname, t_gn_sym) %>%
  magrittr::set_colnames(c("source", "target"))
View(drugGeneEdges)

nervous_system_edges <- nervousSystemDiseases %>%
  dplyr::select(Description, shared_symbol) %>%
  magrittr::set_colnames(c("source", "target")) %>%
  dplyr::bind_rows(drugGeneEdges)
View(nervous_system_edges)
```

#### Nodes

- The key information we need in the nodes table includes:

- **id**: The name of the node
- **label**: The label of the node (same as **id** in most cases)
- **type**: the type of node, this is because we want to color them differently in the network. 

```{r}
nodeType_drug <- data.frame(
  id = drugGeneEdges$source %>% unique,
  type = "drug"
)

nodeType_gene <- data.frame(
  id = c(unique(drugGeneEdges$target), nervousSystemDiseases$shared_symbol),
  type = "gene"
)

nodeType_indication <- data.frame(
  id =unique(nervousSystemDiseases$Description),
  type = "disease"
)

nodes <- bind_rows(
  nodeType_drug,
  nodeType_gene,
  nodeType_indication
) %>%
  dplyr::mutate(label = case_when(
    type == "drug" | type == "disease" ~ id,
    TRUE ~ ""
  ))
View(nodes)
```

#### Export tables

```{r}
nodes %>% readr::write_csv(here("output", "indications_nervous_system_nodes3.csv"))

nervous_system_edges %>%
  readr::write_csv(here("output", "indications_nervous_system_edges3.csv"))
```

### Subnetwork - Only drugs to genes


```{r}
nodes_drugsGenes <- nodes %>%
  dplyr::filter(type %in% c("drug", "gene"))

nodes_drugsGenes$type %>% unique
dim(nodes_drugsGenes)
View(nodes_drugsGenes)

edges_drugsGenes <- nervous_system_edges %>%
  dplyr::filter(source %in% nodes_drugsGenes$id &
                  target %in% nodes_drugsGenes$id)

nodes_drugsGenes  %>%
  readr::write_csv(here("output", "nodes_drugsGenes.csv"))
edges_drugsGenes %>%
  readr::write_csv(here("output", "edges_drugsGenes.csv"))
```

### Subnetwork - Only diseases and genes

```{r}
nodes_diseasesGenes <- nodes %>%
  dplyr::filter(type %in% c("disease", "gene"))

nodes_diseasesGenes$type %>% unique
dim(nodes_diseasesGenes)
View(nodes_diseasesGenes)

edges_diseasesGenes <- nervous_system_edges %>%
  dplyr::filter(source %in% nodes_diseasesGenes$id &
                  target %in% nodes_diseasesGenes$id)

nodes_diseasesGenes  %>%
  readr::write_csv(here("output", "nodes_diseasesGenes.csv"))
edges_diseasesGenes %>%
  readr::write_csv(here("output", "edges_diseasesGenes.csv"))
```

