---
title: "Scratch: Create internal CMap-like database from NESDA data and compare Score approaches"
author: "Nhi Hin"
date: "2023-05-03"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Load Packages

```{r message=FALSE, warning=FALSE}
library(Seurat)
library(dplyr)
library(magrittr)
library(readr)
library(tibble)
library(ggplot2)
library(SingleCellExperiment)
library(scater)
library(here)
library(rhdf5)
library(signatureSearch)
library(signatureSearchData)
library(HDF5Array)
library(ExperimentHub)
library(cowplot)
theme_set(theme_cowplot())
```

## Output Directory

```{r}
output_dir <- here("output", "_scratch_nesda_Rmd")

ifelse(!dir.exists(output_dir),
       dir.create(output_dir), 
       FALSE)
```


## Objective

1. Load NESDA data and get it into a form where we can use it for our internal database of CMap-like datasets, and also in a form where we can use the signatures (significant Somamers as determined previously by Idit) as queries. 

2. Answer the question of which scoring mechanism is best. To test this, we will compare 3 different scoring methods. 

3. Export out the full NESDA database as h5 files ready for use with other datasets. 

The results of this workflow will be used to inform the `workflow_prepare_proteomics_query.Rmd` and `workflow_prepare_proteomics_db.Rmd` notebooks. 

## Load Helper functions

```{r}
source(here("code", "helper_functions.R"))
```


## Import Data foe Databases

- Updated to read in the full datasets for each cohort. 

```{r}
nesda_path <- file.path("/mnt/atlas/work/general/external/studies/NESDA")

nesda_names <- list.files(nesda_path,
                          pattern = "*_full.csv",
                          full.names = FALSE) %>%
  gsub(x = .,
       pattern = ".csv|NESDA_per_somamer_meds_",
       replacement = "")

nesdata <- list.files(nesda_path,
                      pattern = "*_full.csv",
                      full.names = TRUE) %>%
  lapply(readr::read_csv) %>%
  set_names(nesda_names)

nesdata %>% lapply(dim)
```

## Import Somamer to Protein Mappings

```{r}
somaMap <- readRDS("/mnt/atlas/general_archive/data/target_selection/cycle_5/data_in/somalogic_annotations/Somalogic-Human_7K_Annotations-Uniprot_Centric.rds")
View(somaMap)
```

## Initial Exploration

### Number of Somamers Per Drugs Per Cohort

```{r}
table_ndrugsPerSomamerPerCohort <- nesdata %>% 
  lapply(function(x){
    x %>% dplyr::group_by(drug) %>%
      dplyr::summarise(n=n())
  }) %>%
  bind_rows(.id = "cohort") 
table_ndrugsPerSomamerPerCohort %>%
  write_csv(file.path(output_dir,
                      "table_ndrugsPerSomamerPerCohort.csv"))
  
table_ndrugsPerCohort <- table_ndrugsPerSomamerPerCohort %>%
  dplyr::group_by(cohort) %>%
  dplyr::summarize(n=n()) 
table_ndrugsPerCohort %>%
  write_csv(file.path(output_dir, "table_ndrugsPerCohort.csv"))
```

### Drug Overlap Between Cohorts

- The majority of drugs are only in the Anxiety-Depression Cohort, or shared between this Cohort and the Healthy and Control cohorts. 
- Also a sizable number of drugs ~15? that are shared between all Cohorts except Bipolar. 
- Bipolar only has one drug. 

```{r}
table_drugsPerCohort <- table_ndrugsPerSomamerPerCohort %>% 
  dplyr::group_by(drug) %>%
  dplyr::summarize(cohort = list(cohort))
View(table_drugsPerCohort)

plot_overlapDrugsPerCohort <- table_drugsPerCohort %>%
  ggplot(aes(x = cohort))+
  geom_bar() +
  ggupset::scale_x_upset()
plot_overlapDrugsPerCohort
plot_overlapDrugsPerCohort %>%
  ggsave(
    file.path(output_dir, "plot_overlapDrugsPerCohort.pdf"),
    plot = .,
    units = "in",
    width = 14,
    height = 8
  )
```

- Get overlaps as a list

```{r}
table_drugsPerCohort2 <- table_ndrugsPerSomamerPerCohort %>% 
  dplyr::group_by(cohort) %>%
  dplyr::summarize(drug = list(drug)) 

drugsPerCohort_l <- table_drugsPerCohort2$drug %>%
  as.list() %>%
  set_names(table_drugsPerCohort2$cohort)

drugsPerCohort_overlap <- overlapGroups(drugsPerCohort_l)
```


### Fold Change of Somamers Between Drugs & Between Cohorts

```{r}
tables_fc_per_somamer <- nesdata %>%
  bind_rows(.id = "cohort") %>%
  dplyr::arrange(Somamer) %>%
  split(x=., f = .$Somamer) %>%
  lapply(function(x){
    x %>%
      dplyr::group_by(drug) %>%
      dplyr::arrange(drug)
  })

tables_fc <- nesdata %>%
  lapply(function(x) {
    x %>%
      dplyr::group_by(drug) %>%
      dplyr::summarise(
        mean_fc = mean(Estimate),
        sd_fc = sd(Estimate),
        min_fc = min(Estimate),
        max_fc = max(Estimate),
        med_fc = median(Estimate)
      )
  })
  
```



## Prepare Matrix for SignatureSearch DB

- Create various versions of the DB:

1. Connectivity Score is `(1 - pv) * Estimate`.
2. Connectivity Score is `Estimate`. This seems a bit fishy though considering that the fold changes are all quite subtle - leaving out p-value information seems not great. At the same time, fold changes should be directly comparable across drugs 
3. Connectivity Score is -log10(p) * Estimate. This is more similar to that of a Volcano Plot.  The potential disadvantage here is that small p-values could skew the overall result more than the v1 method. 

### All Cohorts

#### Version 1 - Rick method

```{r}
names(nesdata) %<>%
  gsub(x = ., pattern = "_full", 
       replacement = "")

all_drug_names <- nesdata %>%
  lapply(function(x){
    x %>%
      split(x=., f = .$drug) %>%
      names()
  })



all_drug_signatures <- list()
for(cohort in names(nesdata)){
  all_drug_signatures[[cohort]] <- list()
  tmp <- nesdata[[cohort]] %>%
    split(x = ., f = .$drug)
  
  for(drug in all_drug_names[[cohort]]){
    all_drug_signatures[[cohort]][[drug]] <- 
      tmp[[drug]] %>%
      dplyr::select(Estimate, pv, Somamer) %>%
      dplyr::mutate(Score = (1 - pv) * Estimate) %>%
      dplyr::select(Somamer, Score) %>%
      tibble::column_to_rownames("Somamer")
  } 
  all_drug_signatures[[cohort]] %<>% 
    bind_cols() %>%
    set_colnames(paste0(all_drug_names[[cohort]], 
                      "__",
                      cohort,
                      "__",
                      "NESDA"))
  
}

all_drug_signatures_v1 <- all_drug_signatures %>%
  bind_cols()


dim(all_drug_signatures_v1)
View(all_drug_signatures_v1)
```

#### Version 2 - FC Only

```{r}
all_drug_signatures <- list()
for(cohort in names(nesdata)){
  all_drug_signatures[[cohort]] <- list()
  tmp <- nesdata[[cohort]] %>%
    split(x = ., f = .$drug)
  
  for(drug in all_drug_names[[cohort]]){
    all_drug_signatures[[cohort]][[drug]] <- 
      tmp[[drug]] %>%
      dplyr::select(Estimate, pv, Somamer) %>%
      dplyr::mutate(Score = Estimate) %>%
      dplyr::select(Somamer, Score) %>%
      tibble::column_to_rownames("Somamer")
  } 
  all_drug_signatures[[cohort]] %<>% bind_cols() %>%
    set_colnames(paste0(all_drug_names[[cohort]], 
                      "__",
                      cohort,
                      "__",
                      "NESDA"))
  
}

all_drug_signatures_v2 <- all_drug_signatures %>%
  bind_cols()


dim(all_drug_signatures_v2)
View(all_drug_signatures_v2)

```

#### Version 3 - Volcano

```{r}
all_drug_signatures <- list()
for(cohort in names(nesdata)){
  all_drug_signatures[[cohort]] <- list()
  tmp <- nesdata[[cohort]] %>%
    split(x = ., f = .$drug)
  
  for(drug in all_drug_names[[cohort]]){
    all_drug_signatures[[cohort]][[drug]] <- 
      tmp[[drug]] %>%
      dplyr::select(Estimate, pv, Somamer) %>%
      dplyr::mutate(Score = -log10(pv)*Estimate) %>%
      dplyr::select(Somamer, Score) %>%
      tibble::column_to_rownames("Somamer")
  } 
  all_drug_signatures[[cohort]] %<>% bind_cols() %>%
    set_colnames(paste0(all_drug_names[[cohort]], 
                      "__",
                      cohort,
                      "__",
                      "NESDA"))
  
}

all_drug_signatures_v3 <- all_drug_signatures %>%
  bind_cols()


dim(all_drug_signatures_v3)
View(all_drug_signatures_v3)

```

#### Create Databases

```{r}
build_custom_db(
  all_drug_signatures_v1, 
  h5file = file.path(output_dir,
                     "all_drug_signatures_v1.h5"))

build_custom_db(
  all_drug_signatures_v2, 
  h5file = file.path(output_dir,
                     "all_drug_signatures_v2.h5"))

build_custom_db(
  all_drug_signatures_v3, 
  h5file = file.path(output_dir,
                     "all_drug_signatures_v3.h5"))

db_paths <- list(
  rick = file.path(output_dir,
                     "all_drug_signatures_v1.h5"),
  fc_only = file.path(output_dir,
                     "all_drug_signatures_v2.h5"),
  volcano = file.path(output_dir,
                     "all_drug_signatures_v3.h5")
)

db_paths %>% lapply(file.exists)

```


## Which Scoring Method is Best?

### Import Signatures & Split to Drug-Per-Dataframe

- `query_up` and `query_down` are lists of lists that contain the drugs to be used in the Up and Down portions of the query, for each drug-response for each cohort(group). 

```{r}
# Queries include significant Somamers only 
# for each group/cohort. 
query_names <- list.files(
  nesda_path,
  pattern = "NESDA_per_somamer_meds",
  full.names = FALSE) %>%
  .[!grepl("_full|_healthy", .)]  %>%
  gsub(x = ., 
       pattern = "NESDA_per_somamer_meds_|.csv",
       replacement = "")

query_signatures <- list.files(
  nesda_path,
  pattern = "NESDA_per_somamer_meds",
  full.names = TRUE) %>%
  .[!grepl("_full|_healthy", .)] %>%
  lapply(readr::read_csv) %>%
  set_names(query_names) 

query_signatures2 <- list()
query_up <- list()
query_down <- list()

for (group in names(query_signatures)){
  query_signatures2[[group]] <- 
    split(x = query_signatures[[group]],
          f = query_signatures[[group]]$drug)
  
  query_signatures2[[group]] %<>% 
    lapply(function(x){
      x %>%
      dplyr::select(Somamer, Estimate, pv) %>%
      dplyr::mutate(Score = (1 - pv) * Estimate) %>%
      dplyr::select(Somamer, Score) %>%
      dplyr::arrange(Score) %>%
      tibble::column_to_rownames("Somamer") 
    })
  
  query_up[[group]] <- query_signatures2[[group]] %>%
    lapply(function(x){
      x %>%
        dplyr::filter(Score > 0) %>%
        rownames()
    })
  
  query_down[[group]] <- query_signatures2[[group]] %>%
    lapply(function(x){
      x %>%
        dplyr::filter(Score < 0) %>%
        rownames()
    })
  
}
```


### Perform Query

- Note: This takes a while to run and frequently causes the R session to crash (not sure why). 

- Running it piecemeal for now (Rick DB, Depression cohort, Anxiety Cohort | FC DB Depression cohort, Anxiety cohort) for testing purposes for subsequent code, will run full thing overnight sometime. 

```{r}
qsig_preps <- list()
queries <- list()
results_queries <- list()

for(db in names(db_paths)) {
  qsig_preps[[db]] <- list()
  queries[[db]] <- list()
  results_queries[[db]] <- list()
  
  for (cohort in names(query_up)) {
    qsig_preps[[db]][[cohort]] <- list()
    queries[[db]][[cohort]] <- list()
    results_queries[[db]][[cohort]]
    
    for (drug in names(query_up[[cohort]])) {
      qsig_preps[[db]][[cohort]][[drug]] <-
        qSig(
          query = list(upset = query_up[[cohort]][[drug]],
                       downset = query_down[[cohort]][[drug]]),
          gess_method = "LINCS",
          refdb = db_paths[[db]]
        )
      
      queries[[db]][[cohort]][[drug]] <-
        gess_lincs(
          qSig = qsig_preps[[db]][[cohort]][[drug]],
          sortby = "NCS",
          tau = FALSE,
          cmp_annot_tb = NULL,
          cmp_name_col = NULL,
          workers = 2
        )
      results_queries[[db]][[cohort]][[drug]] <- 
     result(queries[[db]][[cohort]][[drug]] )
    } #End drug loop
  } #End cohort loop
} #End DB loop
  
queries %>% saveRDS(file.path(output_dir, "queries_depression_rick.rds"))
results_queries %>% saveRDS(file.path(output_dir, "results_queries_depression_rick.rds"))
```
### Load Saved Results

```{r}
queries <- readRDS(file.path(output_dir, "queries.rds"))
results_queries <- readRDS(file.path(output_dir, "results_queries.rds"))
```


### Compare Rankings

- NCS (Normalized Connectivity Score) is used to rank the results dataframe of each query. 
- If we are performing a query with Drug A (from group Control) against the database, we will pull out the Rank (based off NCS) of Drug A for all cohorts in the query-results dataframe. 
- The corresponding ranks of each drug will be summarized in a table and the 3 methods can be compared by summing the ranks of all drugs. The method with the smallest score should be the best one. 

```{r}
cs_ranks <- list()
rank_results <- list()

for (scoringmethod in names(results_queries)) {
  cs_ranks[[scoringmethod]] <- list()
  
  for (cohort in names(results_queries[[scoringmethod]])) {
    cs_ranks[[scoringmethod]][[cohort]] <- list()
    
    for (d in names(results_queries[[scoringmethod]][[cohort]])) {
      cs_ranks[[scoringmethod]][[cohort]][[d]] <-
        results_queries[[scoringmethod]][[cohort]][[d]] %>%
        #dplyr::filter(cell == cohort) %>%
        dplyr::arrange(desc(NCS)) %>%
        dplyr::mutate(NCS_rounded = signif(NCS, digits = 3)) %>%
        dplyr::mutate(rank = base::rank(-(NCS_rounded), 
                                        ties.method = "min")) %>%
        #dplyr::mutate(rank = 1:nrow(.)) %>%
        dplyr::filter(pert == d) %>%
        dplyr::select(cell, rank)
    }
    
    cs_ranks[[scoringmethod]][[cohort]] %<>%
      bind_rows(.id = "drug") %>%
      dplyr::arrange(drug, cell)
    
    # cs_ranks[[scoringmethod]][[cohort]] %<>%
    #   unlist() %>%
    #   as.data.frame() %>%
    #   set_colnames("rank")
    

  }
}

table(cs_ranks$rick$Anxiety$cell == cs_ranks$fc_only$Anxiety$cell)

rank_res <- list()
for(cohort in names(query_up)){
  rank_res[[cohort]] <- cs_ranks %>%
    lapply(function(x){
      x[[cohort]]
    }) %>%
    # purrr::reduce(function(x){
    #   x %>% full_join(by = c("drug", "cell"))
    # }) %>%
    bind_cols() %>%
  dplyr::select(drug...1, cell...2, starts_with("rank")) %>%
    set_colnames(c("drug", "group", names(results_queries)))
    #set_colnames(names(results_queries))
}
```
```{r}
rank_res %>% saveRDS(file.path(output_dir, "rank_res.rds"))
```

```{r}
rank_res$Anxiety %>% View
```


### Boxplots 


```{r}
rank_boxplots <- list()

for(group in names(rank_res)) {
  rank_boxplots[[group]] <-
    rank_res[[group]] %>%
    melt() %>%
    dplyr::rename(method = variable,
                  rank = value) %>%
    ggplot(aes(x = rank, y = method)) +
    geom_violin(alpha = 0.3, aes(fill = method)) +
    scale_fill_manual(values = pals::glasbey(n = 3)) +
    geom_boxplot(width = 0.2, fill = "white") +
    facet_wrap(~ group) +
    ggtitle(paste0("Query: Drug signatures (significant Somamers per drug) from ", group, " group\n", "Database: Drug signatures from all groups")) +
    labs (x = "Rank of the same drug in each group of database (ranked on NCS (to 3 s.f.)",
          y = "Method of calculating score for somamers in each drug signature in database")
  
  rank_boxplots[[group]] %>%
    ggsave(filename = file.path(output_dir, 
                                paste0("rank_boxplot_",
                                group, ".pdf")),
           plot = ., 
           units = "in",
           width = 13, height = 9)
}

rank_boxplots$Control


```

### Is the same drug significant?

```{r}
sig_res <- list()
for(cohort in names(query_up)){
  sig_res[[cohort]] <- cs_ranks %>%
    lapply(function(x){
      x[[cohort]]
    }) 
}

sig_res <- list()
for (scoringmethod in names(results_queries)) {
  sig_res[[scoringmethod]] <- list()
  
  for (cohort in names(results_queries[[scoringmethod]])) {
    sig_res[[scoringmethod]][[cohort]] <- list()
    
    for (d in names(results_queries[[scoringmethod]][[cohort]])) {
      sig_res[[scoringmethod]][[cohort]][[d]] <-
        results_queries[[scoringmethod]][[cohort]][[d]] %>%
        #dplyr::filter(cell == cohort) %>%
        dplyr::arrange(desc(NCS)) %>%
        dplyr::mutate(NCS_rounded = signif(NCS, digits = 3)) %>%
        dplyr::mutate(rank = base::rank(-(NCS_rounded), 
                                        ties.method = "min")) %>%
        #dplyr::mutate(rank = 1:nrow(.)) %>%
        dplyr::filter(pert == d) %>%
        dplyr::select(cell, NCS, starts_with("WTCS"))
    }
    
    sig_res[[scoringmethod]][[cohort]] %<>%
      bind_rows(.id = "drug") %>%
      dplyr::arrange(drug, cell)
  }
}


rank_res <- list()
for(cohort in names(query_up)){
  rank_res[[cohort]] <- cs_ranks %>%
    lapply(function(x){
      x[[cohort]]
    }) %>%
    # purrr::reduce(function(x){
    #   x %>% full_join(by = c("drug", "cell"))
    # }) %>%
    bind_cols() %>%
  dplyr::select(drug...1, cell...2, starts_with("rank")) %>%
    set_colnames(c("drug", "group", names(results_queries)))
    #set_colnames(names(results_queries))
}
```


### Conclusion
- I couldn't see any clear differences between the rank distributions for each of the 3 methods (Rick, fc only, volcano). They showed fairly similar performance. For cohorts which seemed more similar (e.g. Anxiety and Control), all performed decently. 
- Weirdly, while we would expect the same drug being queries to come up as the top ranked drug, this does not happen much at all. 

