---
title: "scratch_liver_PPF"
author: "Nhi Hin"
date: "2023-02-20"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE}
library(Seurat)
library(dplyr)
library(magrittr)
library(readr)
library(tibble)
library(ggplot2)
library(SingleCellExperiment)
library(scater)
library(here)
library(rhdf5)
library(signatureSearch)
library(signatureSearchData)
library(HDF5Array)
library(ExperimentHub)
```


## Summary

In order to do more relevant testing, I think it would be a good idea to use one of our own datasets as a query dataset. 

Will prepare liver dataset for such a purpose. 

Needs to be run on R 4.2.2, otherwise signatureSearch functions don't work properly. 

## Import Data

```{r}
seurat_path <- "/mnt/home/alkahest.com/dleone/Documents/Analyses/alkahest/sci_app/liver/2023-01-17_PPF/seurat/2023-01-17_filtered.alkahest.main.obj.Rds"

seurat_liver <- readRDS(seurat_path)
```

## Load Helper functions

- The helper functions basically abstract out what was done in the `scratch_6021.Rmd` markdown, but generalising out to other datasets too. It seems to work out of box for liver data although as this data is quite large it takes a while to run. 

```{r}
source(here("code", "helper_functions.R"))
```

## Run Analysis

```{r}
# Note first step takes ~20 mins to run + up to 150Gb in memory 
# depending on dataset due to filtering step
sce <- convertSeuratToSCE(seurat_liver)
gc()

pseudobulk_mat <- aggregateSCEIntoPseudobulkMatrix(sce)
pseudobulk_mat_l <- splitAggregatedMatrixPerCelltype(pseudobulk_mat)
zscores_l <- splitPerCelltypeAggregatedMatrixByGroup(sce, pseudobulk_mat_l)
```

Save data

```{r eval=FALSE}
zscores_l %>% saveRDS(here("data", "zscores_l_liver.rds"))

zscores_l <- readRDS(here("data", "zscores_l_liver.rds"))
```

## Convert mouse ensembl ID to human entrez ID

- Because CMap is in the form of human entrez IDs, we will need to convert the mouse Ensembl IDs in order to do queries. 

- There are many:1 mappings going from mouse to human. In this case we will use the gene expression value with the maximum absolute value (regardless of sign). 

```{r}
# takes too long to run for testing purposes, expect
# 1-2 mins per cell type
# zscores_entrez <- zscores_l %>%
#   lapply(function(x){
#     x %>% lapply(function(y){
#       df <- y %>% 
#         as.data.frame %>%
#         set_colnames(c("zscore")) %>%
#         convertMouseEnsemblToHumanEntrezID(
#           functionForReplicates = "absmax")
#       ls <- f$zscore %>%
#         magrittr::set_names(rownames(df))
#         return(ls)
#     })
#   })
# zscores_df %>% saveRDS(here("zscores_df.rds"))

# Just running for hepatocyte
zscores_entrez <- list()
zscores_entrez$hepatocyte <- zscores_l$hepatocyte %>%
  lapply(function(y){
    df <- y %>% 
         as.data.frame %>%
         set_colnames(c("zscore")) %>%
         convertMouseEnsemblToHumanEntrezID(
           functionForReplicates = "absmax")
    ls <- df$zscore %>%
      magrittr::set_names(rownames(df))
    return(ls)
  })

zscores_entrez$hepatocyte$Vehicle_NC_0hr %>% head

# Try running for endothelial too
zscores_entrez$sinusoidal_endothelial <- zscores_l$`sinusoidal endothelial` %>%
  lapply(function(y){
    df <- y %>% 
         as.data.frame %>%
         set_colnames(c("zscore")) %>%
         convertMouseEnsemblToHumanEntrezID(
           functionForReplicates = "absmax")
    ls <- df$zscore %>%
      magrittr::set_names(rownames(df))
    return(ls)
  })
  
```


## Calculate difference of z-scores for a few relevant comparisons 

- This will be done manually until I figure out how to abstract it out to a function. 

- The idea here is to get a score corresponding to how DE each gene is in a particular comparison. I think we will probably use an actual DE metric for this eventually (for any experiments with n=3), but I can see this method being OK for any experiments with less than 3 samples. 

### Hepatocyte

```{r}
zscores_l$hepatocyte %>% names

hepatocyte_diffs <- list()
hepatocyte_diffs$PPF_vs_Veh_NC_0hr <- zscores_entrez$hepatocyte$`Pl-47615_NC_0hr` - zscores_entrez$hepatocyte$Vehicle_NC_0hr

hepatocyte_diffs$PPF_vs_Veh_NC_0hr %>% summary()

hepatocyte_diffs$PPF_vs_Veh_HFD_0hr <- zscores_entrez$hepatocyte$`Pl-47615_HFD_0hr` - zscores_entrez$hepatocyte$Vehicle_HFD_0hr

hepatocyte_diffs$PPF_vs_Veh_HFD_0hr %>% summary()
```

### Sinusoidal Endothelial

```{r}
se_diffs <- list()

se_diffs$PPF_vs_Veh_NC_0hr <- zscores_entrez$sinusoidal_endothelial$`Pl-47615_NC_0hr` - zscores_entrez$sinusoidal_endothelial$Vehicle_NC_0hr

se_diffs$PPF_vs_Veh_NC_0hr %>% summary()

se_diffs$PPF_vs_Veh_HFD_0hr <- zscores_entrez$sinusoidal_endothelial$`Pl-47615_HFD_0hr` - zscores_entrez$sinusoidal_endothelial$Vehicle_HFD_0hr

se_diffs$PPF_vs_Veh_HFD_0hr %>% summary()
```


## Get top and bottom 150 genes based on ranking of difference of z-scores

```{r}
# upset <- head(names(query[order(-query)]), 150)
#downset <- tail(names(query[order(-query)]), 150)

upsets <- list()
downsets <- list()

upsets$hepatocyte <- list()
downsets$hepatocyte <- list()

upsets$sinusoidal_endothelial <- list()
downsets$sinusoidal_endothelial <- list()


for(comparison in names(hepatocyte_diffs)){
  upsets$hepatocyte[[comparison]] <- head(names(hepatocyte_diffs[[comparison]][order(-hepatocyte_diffs[[comparison]])]), 150)
  
  downsets$hepatocyte[[comparison]] <- tail(names(hepatocyte_diffs[[comparison]][order(-hepatocyte_diffs[[comparison]])]), 150)
}

for(comparison in names(se_diffs)){
  upsets$sinusoidal_endothelial[[comparison]] <- head(names(se_diffs[[comparison]][order(-se_diffs[[comparison]])]), 150)
  
  downsets$sinusoidal_endothelial[[comparison]] <- tail(names(se_diffs[[comparison]][order(-se_diffs[[comparison]])]), 150)
}

```


## Load test CNS internal data base

The generation of the internal database (h5 file) was done in the `scratch_signatureSearch.Rmd` notebook. 

This database only contains the CNS data, subsetted to the PPF and Vehicle samples only (so not young and aged samples; all samples are aged). This is just for testing purposes. 

```{r}
custom_db_path <- here("output","db_6021_signatureSearch.h5")

custom_db_6021 <- SummarizedExperiment(
  HDF5Array(custom_db_path,
            name = "assay")
)

rownames(custom_db_6021) <- HDF5Array(custom_db_path, 
                                      name = "rownames")

colnames(custom_db_6021) <- HDF5Array(custom_db_path,
                                      name = "colnames")

custom_db_6021
```

## Perform query of liver data against CNS data

```{r}
qsigs_hepatocyte <- list()
query_against_customdb_hepatocyte <- list()
for(comparison in names(hepatocyte_diffs)){
  qsigs_hepatocyte[[comparison]] <- qSig(
    query = list(upset = upsets$hepatocyte[[comparison]],
                 downset = downsets$hepatocyte[[comparison]]),
    gess_method = "LINCS",
    refdb = custom_db_path
  )
  
  query_against_customdb_hepatocyte[[comparison]] <- 
    gess_lincs(qSig = qsigs[[comparison]],
               sortby = "NCS",
                  tau = FALSE,
                  workers = 2)
}

qsigs_se <- list()
query_against_customdb_se <- list()
for(comparison in names(se_diffs)){
  qsigs_se[[comparison]] <- qSig(
    query = list(upset = upsets$sinusoidal_endothelial[[comparison]],
                 downset = downsets$sinusoidal_endothelial[[comparison]]),
    gess_method = "LINCS",
    refdb = custom_db_path
  )
  
  query_against_customdb_se[[comparison]] <- 
    gess_lincs(qSig = qsigs_se[[comparison]],
               sortby = "NCS",
                  tau = FALSE,
                  workers = 2)
}

View(result(query_against_customdb_se
            $PPF_vs_Veh_HFD_0hr))
```

- Overall testing two different cell types (hepatocyte; sinusoidal endothelial) against the CNS data, some positive associations, all connectivity scores 0.4 or less, none negative. indicates moderate similarity?

## Query LINCS2

```{r}
lincs2 <- readRDS(here("data", "lincs2.rds"))
```


```{r}
qsigs_hepatocyte2 <- list()
query_against_customdb_hepatocyte2 <- list()
for(comparison in names(hepatocyte_diffs)){
  qsigs_hepatocyte2[[comparison]] <- qSig(
    query = list(upset = upsets$hepatocyte[[comparison]],
                 downset = downsets$hepatocyte[[comparison]]),
    gess_method = "LINCS",
    refdb = lincs2
  )
  
  query_against_customdb_hepatocyte2[[comparison]] <- 
    gess_lincs(qSig = qsigs_hepatocyte2[[comparison]],
               sortby = "NCS",
                  tau = FALSE, 
               cmp_annot_tb = lincs_pert_info2, 
  by = "pert_id", 
  cmp_name_col = "pert_iname",
                  workers = 4)
}
```

#### Functional enrichment analysis

```{r}
drugs_hepatocyte <-query_against_customdb_hepatocyte2$PPF_vs_Veh_NC_0hr %>%
  result %>%
  as.data.frame %>%
  dplyr::filter(WTCS_FDR < 0.05) %>%
  dplyr::pull(pert_iname) %>%
  unique
# Gene ontology
go_tea <- tsea_dup_hyperG(drugs = drugs_hepatocyte,
                          universe="Default",
                          type = "GO",
                          ont = "MF",
                          pvalueCutoff = 0.05,
                          pAdjustMethod = "BH",
                          qvalueCutoff = 0.1,
                          minGSSize = 10,
                          maxGSSize = 500)

result(go_tea) %>% View

dtnetplot(drugs = drugs(go_tea),
          set = "GO:0004674",
          ont = "MF")


```

- KEGG

```{r}
kegg_tea  <- tsea_dup_hyperG(drugs = drugs_hepatocyte,
                          universe="Default",
                          type = "KEGG",
                          pvalueCutoff = 0.5,
                          pAdjustMethod = "BH",
                          qvalueCutoff = 0.1,
                          minGSSize = 10,
                          maxGSSize = 500)

result(kegg_tea) %>% View()
```

- Reactome

```{r}
reactome_tea  <- tsea_dup_hyperG(drugs = drugs_hepatocyte,
                          type = "Reactome",
                          pvalueCutoff = 0.5,
                          pAdjustMethod = "BH",
                          qvalueCutoff = 0.5,
                          readable=TRUE)

result(reactome_tea) %>% View
```

