---
title: "scratch_drugbank"
author: "Nhi Hin"
date: "2023-03-09"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Load Packages
```{r}
library(Seurat)
library(dplyr)
library(magrittr)
library(readr)
library(tibble)
library(ggplot2)
library(SingleCellExperiment)
library(scater)
library(here)
library(rhdf5)
library(signatureSearch)
library(signatureSearchData)
library(HDF5Array)
library(ExperimentHub)
library(fuzzyjoin)
```


## Introduction

- The drug => indication metadata in CMap as provided by *signatureSearch* is quite lacking. Most drugs don't have indications provided. 

- We have a professional account to access DrugBank database. The purpose of this notebook is to explore the info available in this database and get it into a form that can interface with *signatureSearch*.

## Import Data

```{r}
drugbank_dir <- "~/Downloads/Alkahest_DrugBank_SQLite"

all_drugbank_tables <-
  list.files(drugbank_dir, 
             pattern = ".csv", 
             full.names = TRUE) %>%
  set_names(gsub(
    x = list.files(drugbank_dir, 
                   pattern = ".csv",
                   full.names = FALSE),
    pattern = ".csv",
    replacement = ""
  )) %>%  lapply(readr::read_csv)
```

## Get indications and join to relevant tables

- Indication_drug

```{r}
all_drugbank_tables$indication_drugs$indication_id %>% unique %>% length

all_drugbank_tables$indication_drugs$drug_id %>% unique %>% length
```


- Structured_indications table contains the `indication_id` (stored in `id`), as well as the `drug_id`. 

```{r}
View(all_drugbank_tables$structured_indications)
all_drugbank_tables$structured_indications$id %>% unique %>% length
all_drugbank_tables$structured_indications$drug_id %>% unique %>% length
```

- Overlap between first two tables:

- There are more indications in the `structured_indications` table, but all the ones in `indication_drugs` overlap with those in `structured_indications`:

```{r}
table(unique(all_drugbank_tables$indication_drugs$indication_id) %in% unique(all_drugbank_tables$structured_indications$id))
```

- The drugs in both tables do not perfectly overlap. 151 drugs are found in `indication_drugs` but NOT `structured_indications`. 2,703 drugs are found in `structured_indications` but NOT in `indication_drugs`.

```{r}
table(unique(all_drugbank_tables$indication_drugs$drug_id) %in% unique(all_drugbank_tables$structured_indications$drug_id))

table(unique(all_drugbank_tables$structured_indications$drug_id) %in% unique(all_drugbank_tables$indication_drugs$drug_id))
```

- Necause of this, we will make a full joined table that contains the unity of all the drug-indication pairs from the above 2 tables

```{r}
structured_indications2 <- all_drugbank_tables$structured_indications %>%
  dplyr::rename(
    indication_id = id
  ) 

indication_drugs2 <- all_drugbank_tables$indication_drugs

indications_drugs_pairs <- full_join(structured_indications2,
                                     indication_drugs2,
                                     by = c("indication_id", "drug_id"))

indications_drugs_pairs %>% dim
indications_drugs_pairs$indication_id %>% unique %>% length
indications_drugs_pairs$drug_id %>% unique %>% length
```



- To figure out which condition the `indication_id` refers to, we also need a table that joins `indication_id` and `condition_id`

```{r}
View(all_drugbank_tables$indication_conditions)
```

- To figure out what the actual names of the conditions are that are referred to by `condition_id`, we need to join a table that includes both `condition_id` and the names of the conditions, which in this case are stored in the `title` column of the following table. The following table's `condition_id` is instead just named `id` but they refer to the same ID. 

```{r}
View(all_drugbank_tables$conditions)
```

- Lastly, the last relevant table is the `drugs` table, which contains both `drug_id` as well as the actual names of the drugs, under `name` column. 

```{r}
View(all_drugbank_tables$drugs)
```

- With the above tables, we will now perform a series of `full_joins` by the relevant IDs. I will also tidy the data (removing special characters & whitespace, renaming some columns, and also removing any duplicate rows). 

```{r}

indications_struc <- indications_drugs_pairs %>%
  dplyr::full_join(all_drugbank_tables$indication_conditions,
                   by = "indication_id") %>%
  dplyr::full_join(all_drugbank_tables$conditions,
                   by = c("condition_id"="id")) %>%
  dplyr::arrange(drug_id) %>%
  dplyr::full_join(all_drugbank_tables$drugs, by = c("drug_id"="id")) %>%
  set_colnames(gsub(
    x = colnames(.),
    pattern = "\\|",
    replacement = ""
  )) %>% 
  dplyr::select(
    drug_id,
    drugbank_id.y, 
    name,
    indication_id,
    title,
    kind,
    off_label,
    otc_use,
    country,
    dose_form,
    relationship.x,
    relationship.y,
    condition_id,
    meddra_id,
    icd10_id,
    snomed_id,
    type,
    description,
    simple_description,
    clinical_description,
    cas_number,
    protein_formula,
    protein_weight,
    investigational,
    approved,
    vet_approved,
    experimental,
    nutraceutical,
    illicit,
    withdrawn,
    starts_with("moldb"),
    synthesis_patent_id,
    biotech_kind
  ) %>%
  # Clean up some columns:
  dplyr::mutate(
    dose_form = gsub(x = dose_form, pattern = "\\|",
                     replacement = ""),
    relationship.x = gsub(x = relationship.x, pattern = "\\|",
                     replacement = ""),
    relationship.y = gsub(x = relationship.y, pattern = "\\|",
                     replacement = ""),
    biotech_kind = gsub(x = biotech_kind, pattern = "\\|",
                     replacement = ""),
    title = gsub(x = title, pattern = ":|_|'|\"|“|”", replacement = "")
  ) %>%
  dplyr::mutate(title = stringr::str_trim(title)) %>%
  dplyr::filter(!is.na(drug_id)) %>%
  dplyr::distinct(.keep_all = TRUE)

View(indications_struc)

# indications_struc <- all_drugbank_tables$structured_indications %>%
#   dplyr::full_join(all_drugbank_tables$indication_conditions,
#                    by = c("id" = "indication_id")) %>%
#   dplyr::full_join(all_drugbank_tables$conditions,
#                    by = c("condition_id" = "id")) %>%
#   dplyr::arrange(drug_id) %>%
#   dplyr::rename(indication_id = id) %>%
#   #dplyr::full_join(all_drugbank_tables$indication_drugs, by )
#   dplyr::full_join(all_drugbank_tables$drugs, by = c("drug_id" = "id")) %>%
#   set_colnames(gsub(
#     x = colnames(.),
#     pattern = "\\|",
#     replacement = ""
#   )) %>%
#   dplyr::select(
#     drug_id,
#     drugbank_id.y, 
#     name,
#     indication_id,
#     title,
#     kind,
#     off_label,
#     otc_use,
#     country,
#     dose_form,
#     relationship,
#     condition_id,
#     meddra_id,
#     icd10_id,
#     snomed_id,
#     type,
#     description,
#     simple_description,
#     clinical_description,
#     cas_number,
#     protein_formula,
#     protein_weight,
#     investigational,
#     approved,
#     vet_approved,
#     experimental,
#     nutraceutical,
#     illicit,
#     withdrawn,
#     starts_with("moldb"),
#     synthesis_patent_id,
#     biotech_kind
#   ) %>%
#   # Clean up some columns:
#   dplyr::mutate(
#     dose_form = gsub(x = dose_form, pattern = "\\|",
#                      replacement = ""),
#     relationship = gsub(x = relationship, pattern = "\\|",
#                      replacement = ""),
#     biotech_kind = gsub(x = biotech_kind, pattern = "\\|",
#                      replacement = ""),
#     title = gsub(x = title, pattern = ":|_|'|\"|“|”", replacement = "")
#   ) %>%
#   dplyr::mutate(title = stringr::str_trim(title)) %>%
#   dplyr::filter(!is.na(drug_id)) %>%
#   dplyr::distinct(.keep_all = TRUE)
# 
# View(indications_struc)

dim(indications_struc)

indications_struc %>% saveRDS(here("output", "indications_struc.rds"))
indications_struc %>% readr::write_csv(here("output", "indications_drugbank.csv"))
```

## Summary of drugs and indications

### Number of Drugs

- Number of drugs in the database

```{r}
nDrugs <- indications_struc$name %>% unique %>% length
nDrugs
```

### Number of Indications

- Number of indications in the database

```{r}
nIndications <- indications_struc$title %>% unique %>% length
nIndications
```

### Number of drugs that DO NOT have indications:

```{r}
nDrugs_noIndic <- indications_struc %>% 
  dplyr::filter(is.na(title)) %>%
  dplyr::pull(name) %>% unique %>% length

nDrugs_noIndic
nDrugs_noIndic/nDrugs
```

74.5% of drugs do not have indications associated with them...? 

### Number of indications per drug

- The mean indications per drug is ~6. Most drugs (75%) have 4 associated indication or less. 

```{r}
indications_struc %>%
  dplyr::filter(!is.na(title)) %>%
  dplyr::group_by(title) %>%
  dplyr::summarise(n=n()) %>%
  .$n %>% summary()
```

- Histogram summary, for the drugs that have indications associated with them. 

```{r}
upTo50 <- indications_struc %>%
  dplyr::filter(!is.na(title)) %>%
  dplyr::group_by(title) %>%
  dplyr::summarise(n=n()) %>% 
  ggplot(aes(x = n)) +
  geom_histogram(color = "black", fill = "white", bins = 50) +
  labs(x = "Number of indications targeted by a drug",
       y = "Frequency") +
  #ggtitle("Number of indications per drug") +
  xlim(c(0,50)) +
  ylim(c(0,1500))

entirePlot <- 
  indications_struc %>%
  dplyr::filter(!is.na(title)) %>%
  dplyr::group_by(title) %>%
  dplyr::summarise(n=n()) %>% 
  ggplot(aes(x = n)) +
  geom_histogram(color = "black", fill = "white", bins = 50) +
  labs(x = "Number of indications targeted by a drug",
       y = "Frequency") 
  #coord_cartesian(xlim = c(0, 50))

cowplot::plot_grid(upTo50, entirePlot)
  
```

## Substring Matching - Extracting the indication from the description

- I noticed that there are many drugs which have the indication included in the description, but not in the indication column itself. 

- So of the 11,498 drugs which have no indication in the `title` column, 5,315 of them have a `description`. 

```{r}
indications_struc %>%
  dplyr::filter(is.na(title)) %>%
  dplyr::filter(!is.na(clinical_description) | 
                  !is.na(simple_description) |
                  !is.na(description)) %>%
  nrow()
```

- A way to address this is using substring matching. First we need to define a list of strings representing indications which we can match against the description columns. 

### Indications

- To keep the terminology consistent, we will use the `title` from the already labelled indications in the dataset. 

```{r}
indications_ls <- indications_struc$title %>% unique %>%
  tolower() %>%
  gsub(x = . , pattern = "\\(.*\\)", replacement = "") %>%
  gsub(x = . , pattern = "\\(|\\)", replacement = "") %>%
  .[order(nchar(.))] %>%
  rev
length(indications_ls)
indications_string <- indications_ls %>%
 paste(., collapse='|')


# OLD CODE - I tried including the DID's indications too, but 
# due to slight differences in wording from the DrugBank database
# it resulted in much redundancy in the results. 
# did <- readr::read_csv(here("data", "13326_2016_110_MOESM1_ESM.csv"), skip=2)
# head(did) %>% View
# 
# indications_string2 <- did$`target/substring` %>% unique %>% 
#   tolower() %>% 
#   gsub(x = ., pattern = ",.*$", replacement = "") %>%
#   gsub(x = . , pattern = "\\(.*\\)", replacement = "") %>%
#   gsub(x = . , pattern = "\\(|\\)", replacement = "") %>%
#   gsub(x = ., pattern = "\\[.*\\]", replacement = "", perl=T) %>%
#   gsub(x = ., pattern = "\\[|\\]", replacement = "", perl=T) %>%
#   stringr::str_squish() %>%
#   unique  %>%
#   paste(., collapse='|' ) 
# indications_string2
# 
# test <- did$`target/substring` %>% unique %>% 
#   tolower() %>% 
#   gsub(x = ., pattern = ",.*$", replacement = "") %>%
#   gsub(x = . , pattern = "\\(.*\\)", replacement = "") %>%
#   gsub(x = . , pattern = "\\(|\\)", replacement = "") %>%
#   gsub(x = ., pattern = "\\[.*\\]", replacement = "", perl=T) %>%
#   gsub(x = ., pattern = "\\[|\\]", replacement = "", perl=T) %>%
#   stringr::str_squish()
# 
# keepTheseCond <- test %>% lapply(function(x){nchar(x)>3}) %>%
#   unlist()
# 
# indications_string2 <- test[keepTheseCond] %>% unique  %>%
#   paste(., collapse='|' ) 
# 
# indications_string_final <- paste0(indications_string, 
#                                    "|", indications_string2)
```

### Descriptions

- Prepare list of descriptions to match against. This is the query text from which we will perform substring matching, prepared from three different description columns within the Drugbank database. 

- To keep things consistent, convert entire descriptions to lower case and remove brackets and anything between brackets. This is the same processing that will be applied to the `title`s themselves. 

```{r}
descs <- indications_struc %>%
  dplyr::mutate(full_description = paste0(clinical_description, 
                                          " ",
                                          simple_description,
                                          " ",
                                          description)) %>%
  dplyr::mutate(full_description = tolower(full_description)) %>%
  dplyr::mutate(full_description = gsub(x = full_description,
                                        pattern = "\\(.*\\)", 
                                        replacement = "")) %>%
  dplyr::mutate(full_description = gsub(x = full_description,
                                        pattern = "\\(|\\)", 
                                        replacement = "") ) %>%
  dplyr::pull(full_description) %>%
  unique %>%
  set_names(paste0("descTempID", 1:length(.)))

# Approx 8800 unique full descriptions. 
length(descs)
descs[[1]]
```

### Substring Matching Step

- The substring matching step is below. The str_extract_all uses regex() so it does take a while to run, about 10 mins or so.
- Look into replacing with https://cran.r-project.org/web/packages/AhoCorasickTrie/AhoCorasickTrie.pdf Aho-Corasick algorithm at some point which is supposed to be the optimal faster algorithm for this. 

```{r eval=FALSE}
library(AhoCorasickTrie)
descToIndications <- AhoCorasickSearchList(
  keywords = indications_ls,
  textList = descs, 
  alphabet = "ascii")

# I think there is a memory leak somewhere bc this uses 
# 100Gb> of RAM which seems weird. 
```


```{r}
library(parallel)
descToIndications <- descs %>%
  mclapply(function(x){
    stringr::str_extract_all(x, 
                         pattern = indications_string)
  })

names(descToIndications) <- names(descs)

descToIndications2 <- descToIndications %>%
  lapply(function(x){
    x[[1]] %>% stringr::str_trim() %>% unique
  })

length(descToIndications2)

descToIndications3 <- descToIndications2 %>%
  lapply(function(x){
    paste0(x, collapse = ",")
  })

descToIndications_df <- data.frame(
  full_description = descs,
  full_description_id = names(descs)
)
descToIndications_df$keywords <- unlist(descToIndications3)
```

### Join back to full data

```{r}
indications_full <- indications_struc %>%  dplyr::mutate(full_description = paste0(clinical_description, 
                                          " ",
                                          simple_description,
                                          " ",
                                          description)) %>%
  dplyr::mutate(full_description = tolower(full_description)) %>%
  dplyr::mutate(full_description = gsub(x = full_description,
                                        pattern = "\\(.*\\)", 
                                        replacement = "")) %>%
  dplyr::mutate(full_description = gsub(x = full_description,
                                        pattern = "\\(|\\)", 
                                        replacement = "") )  %>%
  dplyr::left_join(descToIndications_df, by = "full_description") %>%
  dplyr::mutate(title = tolower(title)) %>%
  dplyr::mutate(title = gsub(x = title,
                                        pattern = "\\(.*\\)", 
                                        replacement = "")) %>%
  dplyr::mutate(title = gsub(x = title,
                                        pattern = "\\(|\\)", 
                                        replacement = "") )  %>%
  dplyr::mutate(newIndication = case_when(
    !is.na(title) ~ title,
    TRUE ~ keywords
  )) %>%
  tidyr::separate_rows(newIndication, sep = ",") %>%
  dplyr::mutate(newIndication = stringr::str_trim(newIndication)) %>%
  dplyr::mutate(newIndication = case_when(
    nchar(newIndication) == 0 ~ NA,
    newIndication == "NA" ~ NA,
    TRUE ~ newIndication
  )) %>%
  dplyr::mutate(hasIndication = !is.na(newIndication))

indications_full$hasIndication %>% table 

dim(indications_full)
indications_full$name %>% unique %>% length
```

OK, so now, the majority of rows appear to have indications associated with them. 


## Tidy up and prepare final drug-indication mappings

```{r}
indications_full_final <- indications_full %>%
  dplyr::select(drug_id, name, cas_number, newIndication) %>%
  dplyr::rename(indication = newIndication) %>%
  dplyr::distinct(name, cas_number, indication,
                  .keep_all=TRUE) %>%
  dplyr::filter(!is.na(indication)) %>%
  dplyr::filter(!is.na(name))
dim(indications_full_final)
indications_full_final$name %>% unique %>% length
indications_full_final$cas_number %>% unique %>% length
indications_full_final$indication %>% unique %>% length
indications_full_final$drug_id %>% unique %>% length

View(indications_full_final)

indications_full_final %>% saveRDS(here("data", "indications_full_final2.rds"))
```

### Import Synonyms

```{r}
indications_full_final_syn <- indications_full_final %>%
  dplyr::left_join(all_drugbank_tables$drug_synonyms %>%
                     dplyr::select(drug_id, synonym),
                   by = "drug_id", multiple = "all")

indications_full_final_syn %>%
  saveRDS(here("data", "indications_full_final_syn.rds"))

```


### Number of indications per drug (again)

- The mean indications per drug is now 4.6.  

```{r}
indications_full_final %>%
  dplyr::filter(!is.na(indication)) %>%
  dplyr::group_by(indication) %>%
  dplyr::summarise(n=n()) %>% 
  .$n %>% summary()
```

- Histogram summary, for the drugs that have indications associated with them. 

```{r}
upTo50_2 <- indications_full_final %>%
  dplyr::filter(!is.na(indication)) %>%
  dplyr::group_by(indication) %>%
  dplyr::summarise(n=n()) %>% 
  ggplot(aes(x = n)) +
  geom_histogram(color = "black", fill = "white", bins = 50) +
  labs(x = "Number of indications targeted by a drug",
       y = "Frequency") +
  #ggtitle("Number of indications per drug") +
  xlim(c(0,50)) +
  ylim(c(0,1500))

entirePlot <- 
  indications_full_final %>%
  dplyr::filter(!is.na(indication)) %>%
  dplyr::group_by(indication) %>%
  dplyr::summarise(n=n()) %>% 
  ggplot(aes(x = n)) +
  geom_histogram(color = "black", fill = "white", bins = 50) +
  labs(x = "Number of indications targeted by a drug",
       y = "Frequency") 
  #coord_cartesian(xlim = c(0, 50))

cowplot::plot_grid(upTo50_2, entirePlot)
  
```


## Try another database (DID)

See https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5223332/ 

```{r}
did <- readr::read_csv(here("data", "13326_2016_110_MOESM1_ESM.csv"), skip=2)
head(did) %>% View

# Number of unique indications. 
did$`target/substring` %>% unique %>% length

#raw_drug_name and PT...6
```

### Extract relevant col

```{r}
did_subset <-
  did %>% dplyr::select(
    PT...30,
    `PT ("name")`,
    ends_with("CAS#"),
    ends_with("substring")
  ) %>%
  dplyr::rename(name = PT...30,
                name2 = `PT ("name")`,
                CAS = `Preferred CAS#`,
                indication = `target/substring`) %>%
  dplyr::mutate(name2 = case_when(
    is.na(name2) ~ name,
    TRUE ~ name2
  )) %>%
  dplyr::mutate(name2 = tolower(name2)) %>%
  dplyr::select(-name) %>%
  dplyr::rename(name = name2)
```

## Worth joining by CAS?

```{r}
cas_joined <- inner_join(did_subset, indications_full_final,
                         by = c("CAS"="cas_number"))%>%
  dplyr::filter(!is.na(CAS))
dim(cas_joined)
View(cas_joined)
```



```{r}
descToIndications_df %<>%
  tidyr::separate_rows(keywords, sep = ",") %>%
  dplyr::mutate(keywords = case_when(
    nchar(keywords) == 0 ~ NA,
    keywords == "NA" ~ NA,
    TRUE ~ keywords
  ))
View(descToIndications_df)
```

- Join back to the original data

```{r}
indications_full <- indications_struc %>%
  dplyr::mutate(full_description = paste0(clinical_description, 
                                          " ",
                                          simple_description,
                                          " ",
                                          description)) %>%
  dplyr::left_join(descToIndications_df, by = "full_description")
  
View(indications_full)
```

```{r}
indications_full %>%
  dplyr::filter(is.na(title)) %>%
    dplyr::mutate(hasIndication = !is.na(keywords)|!is.na(title)) %>% View()

.$hasIndication %>% table
```

## Old Code


```{r}
indications_struc$title

indications_struc2 <- indications_struc %>%
  dplyr::mutate(indication_from_desc = stringr::str_extract_all(string = clinical_description, pattern = indications_string))
```


## Try another database (DID)

See https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5223332/ 

```{r}
did <- readr::read_csv(here("data", "13326_2016_110_MOESM1_ESM.csv"), skip=2)
head(did) %>% View

did$`target/substring` %>% unique %>% length

#raw_drug_name and PT...6
```

### Fuzzy matching on drug names

- Attempt to join the DID database to the Drugbank ones through fuzzy string matching

```{r}
did_test <-
  did %>% dplyr::select(
    `raw drug name`,
    PT...6,
    PT...30,
    ends_with("CAS#"),
    `MESH ID`,
    raw,
    starts_with("entire"),
    ends_with("substring")
  ) %>%
  dplyr::rename(name = PT...30,
                CAS = `Preferred CAS#`)
```

- NOTE: Takes a while to run. 

```{r}
fuz_test <- stringdist_join(indications_struc, did_test[1:500,], 
                by='name', #match based on team
                mode='left', #use left join
                method = "jw", #use jw distance metric
                max_dist=99, 
                distance_col='dist') %>%
  group_by(name.x) %>%
  slice_min(order_by=dist, n=1)
```

### CAS Number

```{r}
indications_struc %>% dplyr::filter(is.na(title)) %>% .$cas_number %>% unique %>% length

indications_struc$cas_number %>% unique %>% length
did_test$CAS %>% unique %>% length

cas_joined <- dplyr::inner_join(indications_struc,
          did_test, by = c("cas_number"="CAS"))

```

