---
title: "scratch_signatureSearch"
author: "Nhi Hin"
date: "2023-02-15"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Summary

- Rather than reimplementing the MATLAB implementation of Connectivity Map's queries into R, it looks like the R package *signatureSearch* does something similar. 

```{r load-packages}
library(signatureSearch)
library(signatureSearchData)
library(ExperimentHub)
library(rhdf5)
library(here)
library(HDF5Array)
```

## Reference Databases

The following databases are available from ExperimentHub:

- CMAP (moderated z-scores)
- CMAP_EXP (Expression values)
- LINCS (moderated z-scores from initial LINCS dataset)
- LINCS2 (moderated z-scores from latest LINCS dataset)


```{r}
eh <- ExperimentHub()
lincs2 <- eh[["EH7297"]]
saveRDS(lincs2, here("data", "lincs2.rds"))
```

- The LINCS2 dataset is moderated z-scores from DE analysis of ~12,000 genes from ~30,000 compound treatments of 58 cell lines. This corresponds to ~136,000 signatures in total. 

- This is the same database used for the Query Tool in the CLUE.io website so it's what we will focus on. 

```{r}
ssd <- query(eh, c("signatureSearchData"))
ssd
```


## Test - Custom Databases (6021)

- Custom database can be built using the `build_custom_db()` function. 

- Input is a matrix or data.frame. Rows are gene IDs (e.g. entrez gene ids). Column names are expected to have structure: (drug)_(cell)_(factor), e.g. sirolimus_endothelial_trt_cp. 

- Values in the matrix/data.frame can include: normalized gene expression counts; log2 fold changes; Z-scores or p-values from DE analysis. 

- Output is a HDF5 file which is where the database will be stored. 

- May as well test this out with the 6021 data we previously tried to prepare for an internal database. 

```{r build_custom_db}
db_6021 <- readr::read_tsv(here("output", "db_entrez_6021.txt")) %>%
  tibble::column_to_rownames("entrezid")

build_custom_db(db_6021, h5file = here("output", "db_6021_signatureSearch.h5"))

custom_db_path <- here("output","db_6021_signatureSearch.h5")

custom_db_6021 <- SummarizedExperiment(
  HDF5Array(custom_db_path,
            name = "assay")
)

rownames(custom_db_6021) <- HDF5Array(custom_db_path, name = "rownames")
colnames(custom_db_6021) <- HDF5Array(custom_db_path, name = "colnames")
custom_db_6021
```

## Test from Vignette

- The toy database is smaller than LINCS(2) so it will be fine for testing the capabilities of the vignette. 

```{r}
db_path <- system.file("extdata", "sample_db.h5", 
                       package = "signatureSearch")

sample_db <- SummarizedExperiment(HDF5Array(db_path, name = "assay"))

rownames(sample_db) <- HDF5Array(db_path, name = "rownames")
colnames(sample_db) <- HDF5Array(db_path, name = "colnames")
```

- The query matrix is simply a one-column matrix, where rows are entrez gene IDs and the column is the particular comparison for which DE was done (e.g. vorinostat-treated vs. control). 

- The top ranked 150 genes and bottom ranked 150 genes (sorted by moderated z-score?) are taken for running CMAP-like and LINCS-like queries (see below). 

```{r}
query_mat <- as.matrix(assay(sample_db[, "vorinostat__SKB__trt_cp"]))
query <- as.numeric(query_mat)
names(query) <- rownames(query_mat)

upset <- head(names(query[order(-query)]), 150)
downset <- tail(names(query[order(-query)]), 150)
```

### Using CMAP search method:

```{r}
qsig_cmap <- qSig(query = list(upset = upset,
                               downset = downset),
                  gess_method = "CMAP", 
                  refdb = db_path)

cmap <- gess_cmap(qSig = qsig_cmap,
                  chunk_size = 5000,
                  workers = 1)
result(cmap) %>% View() 
```

### Using new LINCS method:

```{r}
qsig_lincs <- qSig(query = list(upset = upset,
                               downset = downset),
                  gess_method = "LINCS", 
                  refdb = db_path)

lincs <- gess_lincs(qSig = qsig_lincs,
                  sortby = "NCS",
                  tau = FALSE,
                  workers = 1)
result(lincs) %>% View() 
```

#### Test on the full LINCS2 data base:

```{r}
qsig_lincs_full <- qSig(query = list(upset = upset,
                               downset = downset),
                  gess_method = "LINCS", 
                  refdb = lincs2)

lincs2_full <- gess_lincs(qSig = qsig_lincs_full,
                  sortby = "NCS",
                  tau = FALSE,
                  workers = 2)
result(lincs2_full) %>% View

# Provide compound annotation table
# Takes ~15 mins
lincs2_with_annot <- gess_lincs(
  qsig_lincs_full,
  tau = TRUE,
  sortby = "NCS",
  workers = 1,
  cmp_annot_tb = lincs_pert_info2, 
  by = "pert_id", 
  cmp_name_col = "pert_iname")
```

#### Test on custom db

```{r}
qsig_custom <- qSig(query = list(upset = upset,
                               downset = downset),
                  gess_method = "LINCS", 
                  refdb = custom_db_path)

lincs2_custom <- gess_lincs(qSig = qsig_custom,
                  sortby = "NCS",
                  tau = FALSE,
                  cmp_annot_tb = NULL,
                  cmp_name_col = NULL, #Need or it fails
                  workers = 2)

result(lincs2_custom) %>% head()
```

### Visualisation

```{r}
qsig_lincs_full <- qSig(query = list(upset = upset,
                               downset = downset),
                  gess_method = "LINCS", 
                  refdb = lincs2)

lincs2_full <- gess_lincs(qSig = qsig_lincs_full,
                  sortby = "NCS",
                  tau = FALSE,
                  workers = 2)

vori_tb <- result(lincs2_full)

drugs_top10 <- unique(result(lincs)$pert)[1:10]

drugs_hdac <- c(
  "panobinostat",
  "mocetinostat",
  "ISOX",
  "scriptaid",
  "entinostat",
  "belinostat",
  "HDAC3-selective",
  "tubastatin-a",
  "tacedinaline",
  "depudecin"
)
drugs = c(drugs_top10, drugs_hdac)
gess_res_vis(vori_tb, drugs = drugs, col = "NCS")

vor_qsig_full <- qSig(
  query = list(upset=upset, downset=downset), 
  gess_method="LINCS", 
  refdb=lincs2)

vori_res_full <- gess_lincs(
  qSig=vor_qsig_full, 
  sortby="NCS", 
  tau=TRUE)

vori_tb <- result(vori_res_full)

drugs_top10 <- unique(result(lincs)$pert)[1:10]

drugs_hdac <- c("panobinostat",
              "mocetinostat",
              "ISOX","scriptaid","entinostat",
      "belinostat","HDAC3-selective","tubastatin-a","tacedinaline","depudecin")
drugs = c(drugs_top10, drugs_hdac)
gess_res_vis(vori_tb, drugs = drugs, col = "NCS")
```

