---
title: "workflow_disgenet"
author: "Nhi Hin"
date: "2023-03-30"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Load Packages

```{r}
library(disgenet2r)
library(Seurat)
library(dplyr)
library(magrittr)
library(readr)
library(tibble)
library(ggplot2)
library(SingleCellExperiment)
library(scater)
library(here)
library(rhdf5)
library(signatureSearch)
library(signatureSearchData)
library(HDF5Array)
library(ExperimentHub)
```

## Project-Specific Variables

```{r}
date_string <- as.character(Sys.Date())
user_notebook <- Sys.info()[["user"]]
output_stem <- here("output")
set.seed(1)
Sys.setenv(DISGENET_API_KEY = "030b2e02ff356ed135085de8c366cbbacb3bf9a4")

output_dir <- file.path(output_stem, project_name, 
                        paste0(date_string, "_",
                               user_notebook))
  
ifelse(!dir.exists(output_dir),
         dir.create(output_dir, recursive = TRUE),
         FALSE)
```

## Summary of this notebook

- Given results table from running CMap through *signatureSearch*, we need a way to interpret the results in terms of enriched indications. 
- In this notebook, we perform over-representation analysis using DisGeNet's disease-gene associations. 

## Import data

### CMap Results (PPF vs Veh, Endothelial cells, vs. HUVEC LINCS2)

See `prepare_6021_data_for_querying_LINCS2.Rmd` for details. 

```{r}
full_table <- readRDS(here("data", "full_table_endo.rds"))
colnames(full_table)
```

## Neaten up table 

- Retain relevant columns
- Separate out gene symbols, one on each row. 

**Note**: Below, here we are focusing on the HUVEC cell line as this is most comparable to endothelial cells from the transcriptlomics data. 

```{r}
full_table_tidy <- full_table %>%
  dplyr::filter(cell == "HUVEC") %>%
  dplyr::select(pert, pert_iname, 
                starts_with("WTC"),
                starts_with("NCS"),
                MOAss,
                t_gn_sym,
                sig)   %>%
  tidyr::separate_rows(t_gn_sym, sep = ";") %>%
  dplyr::mutate(t_gn_sym =  stringr::str_trim(t_gn_sym)) 
nrow(full_table_tidy)

head(full_table_tidy)
```

- Number of gene targets per drug:

```{r}
full_table_tidy %>%
  dplyr::distinct(pert_iname, t_gn_sym) %>%
  dplyr::group_by(pert_iname) %>%
  dplyr::summarise(n=n()) %>%
  dplyr::pull(n) %>%
  hist(breaks = 50)
```

-Benoit: what kinds of drugs have 20> targets

```{r}
drugswith20targets <- full_table_tidy %>%
  dplyr::distinct(pert_iname, t_gn_sym) %>%
  dplyr::group_by(pert_iname) %>%
  dplyr::summarise(n=n()) %>%
  dplyr::filter(n >= 20)


```

## Filter sig results, extract gene symbols

```{r}
sig_res <- full_table_tidy %>%
  dplyr::filter(sig == TRUE) 
  

sig_res %>%
  dplyr::distinct(pert_iname, t_gn_sym) %>% 
  dplyr::group_by(pert_iname) %>%
  dplyr::summarise(n=n()) %>% .$n %>%
  hist(breaks=50)

  sig_res <- sig_res%>% .$t_gn_sym %>% unique()
sig_res
length(sig_res)
```

